---
output:
  pdf_document: default
  html_document: default
---
# Reproducible Research

## Loading and Processing the data

### 1. Downloading and loading the data into R workspace.

```{r, echo=TRUE}

activitydata <- read.csv("activity.csv", header = T)

dim(activitydata)

head(activitydata)

str(activitydata)

activitydatacopy <- activitydata
```

This shows that the date coulmn of our data is not in date format, so we have to convert it to the date format.

```{r, echo=TRUE}
activitydata$date <- as.Date(activitydata$date, "%Y-%m-%d")
```
Now the data is loaded and processed for our analysis.

### 2. What is mean total number of steps taken per day and the Histogram represntation.

Code for calculating the total number of steps per day.

```{r, echo=TRUE}
aggregatesteps <- tapply(activitydata$steps, activitydata$date, FUN = sum, na.rm = T)
hist(aggregatesteps, breaks = 19, main = "Histogram of total number of steps perday", xlab = "Total number of steps per day", col = "green")

```

### 3. Code for calculating the mean and median of the total number of steps taken per day.

```{r, echo=TRUE}
meansteps <- mean(aggregatesteps)
print(meansteps)

mediansteps <- median(aggregatesteps)
print(mediansteps)
```

### 4. Time series plot of the average number of steps taken.

```{r, echo=TRUE}

avgsteps_perday <- tapply(activitydata$steps, activitydata$interval, FUN = mean, na.rm = T)
plot(avgsteps_perday, type = "l", ylab = "average number of steps", xlab = "5 minutes interval", main = "Time series plot")
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo=TRUE}
avgsteps_perday[which.max(avgsteps_perday)]
                
```
### 6. Code to describe and show a strategy for imputing missing data.
Stategy used: Replacing the NAs with integer 0.

```{r, echo=TRUE}
activitydata1 <- activitydata ## A copy of the orignal data is made so as to make two histograms with orignal and imputed NAs for comparision.

activitydata1$steps[is.na(activitydata1$steps)] <- 0

head(activitydata1, 20) ## To check whether this worked or not.
```

### 7. Histogram of the total number of steps taken each day after missing values are imputed.

```{r, echo=TRUE}
aggregatesteps <- tapply(activitydata$steps, activitydata$date, FUN = sum, na.rm = T) ## This is the orignal data frame with NAs.


aggregatesteps1 <- tapply(activitydata1$steps, activitydata1$date, FUN = sum) ## Notice here that I have not used the na.rm argument since I am using the data frame whose NAs are imputed to zero.


par(mfrow = c(1, 2))

hist(aggregatesteps, breaks=19, main = "Orignal data without NA imputation", xlab = "total steps taken per day")

hist(aggregatesteps1, breaks=19, main = "Data with NAs imputed to 0", xlab = "total steps taken per day")

```
### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.

For this, first I have to subset my original data into weekdays and weekends.

```{r, echo=TRUE}

# Converting the dates in the actual data to Weekdays and then subset into weekdays and weekends.

activitydatacopy$day <- strftime(activitydatacopy$date, "%A") # Now the data has names of weekdays instead of dates.

activityweekday <- activitydatacopy[activitydatacopy$day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), ]

activityweekend <-  activitydatacopy[activitydatacopy$day %in% c("Saturday", "Sunday"), ]

```

###  Ploting the panel graph for the two subsets.

```{r, echo=TRUE}

dim(activityweekday)
dim(activityweekend)
# This is just to check that my subsets are correct.

weekdayavg <- tapply(activityweekday$steps, activityweekday$interval, FUN = mean, na.rm = T)

weekendavg <- tapply(activityweekend$steps, activityweekend$interval, FUN = mean, na.rm = T)

par(mfrow = c(1, 2))

plot(weekdayavg, type = "l", ylab = "average number of steps", xlab = "5 minutes interval", main = "Weekday Time series plot")

plot(weekendavg, type = "l", ylab = "average number of steps", xlab = "5 minutes interval", main = "Weekend Time series plot")
```


### 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report.


















